\documentclass{article}

\usepackage{iclr2025_conference,times}
\input{math_commands.tex}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algpseudocode}
% \usepackage[parfill]{parskip}

% Replace ICLR camera-ready publication banner with course-project notice.
\makeatletter
\def\@noticestring{CSC415 course project report}
\makeatother

\title{Inference-Time Adaptive Execution for Transformer Action Chunking in Robotic Manipulation\\\vspace{0.35em}\large CSC415 course project (final report)}
\author{Zuhayr Ahmed, Charles Cheung, Lawrence Fu}
\date{April 2026}

\iclrfinalcopy

\begin{document}

\maketitle
% Override ICLR running header text for course-project submission.
\lhead{CSC415 course project report}

\begin{abstract}
Action chunking improves long-horizon control, but fixed chunk execution can react too slowly near contact events and can introduce boundary inconsistency. We train an ACT-style chunk policy with offline imitation and focus on an inference-time execution layer that decides when to replan. We compare fixed execute-all and receding-horizon execution against adaptive overlap-disagreement commits, plus decide-at-start and learned-commit variants, on RoboMimic Lift, Can, and Square with low-dimensional observations. On Lift, dynamic overlap reaches 1.00 success while reducing replans per episode from 399.0 (receding) to 86.65; Can is mixed and Square remains challenging, showing where adaptive execution helps and where stronger policy learning is still needed.
\end{abstract}

\section{Introduction}

Long-horizon robotic manipulation benefits from temporal abstraction, but policies must remain reactive near contact-rich decision points (e.g., grasping/insertion). Transformer-based action chunking predicts $k$ future actions to reduce decision frequency and compounding error, but most methods use a fixed $k$ (e.g., $k \approx 100$), which can reduce reactivity and introduce boundary instability. In particular, chunk switches can cause discontinuities or chunk-boundary inconsistency/jitter in real-time execution \cite{real_time_chunking,bidirect}.

\textbf{Motivation for disagreement-based execution.} Prior work shows overlap consistency is important when chunk predictions are refreshed under real-time constraints \cite{real_time_chunking}. In our setting, this suggests a simple operational test: if the newly predicted overlap disagrees with the queued overlap, continuing to commit the old queue is likely harmful. We therefore use overlap disagreement as a trigger to preempt and replan, while leaving the trained policy weights unchanged.

\textbf{Imitation training, RL-style evaluation.} We fit the chunking policy from a \emph{fixed offline dataset} of expert trajectories (behavioral cloning), then evaluate with RL-style rollout metrics in simulation: return, success, and trajectory statistics. The environment remains a standard finite-horizon control problem; our contribution is the execution-time rule on top of the learned policy.

\textbf{Objectives.} We (1) implement adaptive commit-length execution using inference-time disagreement/uncertainty triggers and optional learned $\hat{m}$, (2) compare dynamic execution against execute-all, receding-horizon, and ACT-style temporal ensembling on the \emph{same} trained chunk policy, and (3) report rollout metrics plus optional Gaussian observation noise and multi-seed retraining for robustness.

\textbf{Scope.} We use RoboMimic with low-dimensional state on Lift, Can, and Square. All three tasks are run under a shared ACT-Strong protocol and matched baseline evaluation pipeline.

\noindent\textbf{Code.} Implementation and reproduction instructions are available at \url{https://github.com/Jojoz23/actdyn}.


\section{Related Work}
\subsection{Transformer-based Sequence Modeling}

Sequence-modeling approaches reframe RL as token prediction over trajectories, using Transformers and self-attention to capture long-range dependencies among states, actions, and returns.

Decision Transformer \cite{chen2021decisiontransformerreinforcementlearning} is an early influential approach in this direction: it frames RL as conditional sequence prediction, models trajectories autoregressively with a causal Transformer conditioned on past states, actions, and a target return.

Trajectory Transformer \cite{janner2021offlinereinforcementlearningbig} extends that work and learns a generative model over environment transitions and full trajectories. States, actions and rewards are discretized and treated as tokens in the sequence model. Instead of directly producing actions, planning is performed by trajectory sampling or a search (beam search) at inference time.

These papers establish Transformers as a viable architecture for RL sequence modeling. This is a different modeling choice than the tabular or function-approximation value-based methods from standard MDP formulations (e.g., approximate $V$ or $Q$ and derive a policy), but policies from either family can still be compared fairly in the \emph{same} simulator by rollout return or success. However, these sequence-modeling approaches mainly focus on planning quality and do not explicitly address real-time execution constraints introduced by action chunking (i.e., when to switch or commit while the robot is already moving) \cite{real_time_chunking}.


\subsection{Action Chunking}

Action chunking provides temporal abstraction by predicting $k$ future actions at once rather than one action at a time, shortening the effective horizon by a factor of $k$, reducing compounding error, and often improving stability.

Action Chunking with Transformers \cite{zhao2023} shows a Transformer-based policy specifically for robotics imitation learning. The policy predicts action chunks which are conditioned on visual observations and robot states, and demonstrates that Action Chunking can effectively model long-horizon robot manipulation performance. 

Reinforcement Learning with Action Chunking \cite{li2025reinforcementlearningactionchunking} formalizes Action Chunking in RL as ``Q-Chunking''. The paper does not treat RL as a sequence modeling problem and still relies on value functions and policy gradient. The policy introduces multi-step action prediction directly into the reinforcement learning objective. The predictions are made in an extended-action-space.

Bidirectional Decoding: Improving Action Chunking via Guided Test-Time Sampling \cite{bidirect} introduces a new algorithm called Bidirectional Decoding to improve the agent's ability to react to unexpected events when using action chunking. This run-time inference algorithm searches for possible predictions and chooses the optimal one based on both alignment with previous decisions and likelihood for future decisions, allowing for both long-term consistency and short-term reactivity, increasing the performance of the base generative algorithm.

Mixture of Horizons in Action Chunking \cite{mixtureofhorizons} proposes a new strategy splitting the action chunk into multiple segments with different horizons, providing both benefits of better long-term planning from longer horizons and increased precision of shorter horizons. 

These papers show that Action Chunking significantly improves long-horizon performance, particularly for environments with sparse rewards or with compounding error from sequential decision making. Nevertheless, many approaches still rely on a fixed horizon $k$ and/or on smoothing strategies (e.g., temporal ensembling) to deal with chunk boundaries, which can reduce reactivity or lead to boundary discontinuities when switching occurs in real time \cite{real_time_chunking}.


\subsection{Our Work}

Most action-chunking policies use a fixed chunk length (e.g., $k \approx 100$), which improves long-horizon control but can reduce reactivity near decision points and cause jitter/inconsistency at chunk boundaries. Prior work on \emph{real-time} chunk execution addresses consistency across overlapping predictions under latency \cite{real_time_chunking}; bidirectional decoding and horizon mixtures address reactivity and multi-scale commitment in complementary ways \cite{bidirect,mixtureofhorizons}. We propose \textbf{adaptive chunk execution} in our setting: predict up to a maximum $k_{\max}$ actions with a Transformer chunk policy, but choose when to replan using inference-time disagreement/uncertainty triggers that preempt inconsistent chunk continuations.

Relative to these lines, our prototype is closest in spirit to latency-aware overlap consistency \cite{real_time_chunking}, but we emphasize a simple \emph{scalar disagreement test} between the queued segment and a fresh prediction to trigger preemption and shorten commits under conflict, without bidirectional search over decoded candidates \cite{bidirect} and without learning an explicit multi-horizon mixture \cite{mixtureofhorizons}. The goal is a lightweight execution layer that reuses the same trained chunk policy while changing only how actions are committed over time.

\section{Methodology}

\subsection{Method}

\paragraph{Policy backbone (ACT-aligned chunk model).}
Our \emph{training} target is the same object as Action Chunking with Transformers \cite{zhao2023}: a Transformer-based policy that outputs a length-$k_{\max}$ sequence of continuous actions from a state observation. We implement this as a \textbf{conditional VAE} over chunks (latent $z$, prior/posterior, reconstruction loss, KL) with observation encoder and action decoder, matching the spirit of ACT-style chunking in low-dimensional RoboMimic. We do \emph{not} claim a new backbone architecture; \textbf{our contribution is inference-time execution}---how predicted chunks are committed, queued, preempted, or blended---holding the trained weights fixed while varying execution modes.

We extend this chunk policy so that at rollout it predicts an action chunk $(a_t,\dots,a_{t+k_{\max}-1})$ (with $k_{\max}$ as a maximum bound). At inference time, we maintain an action queue of already-committed actions and decide the next commit length (how many actions to keep executing before replanning) using inference-time triggers. Unless stated otherwise, ``dynamic'' in this report refers to the overlap-disagreement rule; decide-at-start rules are ablations. Crucially, \emph{this commit-length decision does not require ground-truth labels}: we use discrepancy measures between overlapping action predictions to detect inconsistent chunk boundary behavior.

\paragraph{Execution-time state and action queue.}
Our executor keeps a FIFO (first-in, first-out) queue of planned actions. When the queue is empty, it plans a new deterministic action chunk from the current observation and chooses an initial commit length based on the evaluation mode (execute-all, receding-horizon, or a dynamic rule).

\paragraph{Primary dynamic rule: overlap disagreement (mid-execution preemption).}
While a queue is active, we repeatedly:
1. Predict a fresh candidate chunk from the current observation.
2. Compute an overlap disagreement score between the candidate chunk and the remaining queued actions over an overlap window of size $w$.
3. If the disagreement score exceeds a threshold $\tau$, we \emph{preempt} the current commit and trigger an immediate replan.

When a preemption happens, we also choose a new planned commit length as a decreasing function of the overlap disagreement score (higher disagreement $\Rightarrow$ shorter commits), clamped to $[k_{\min},k_{\max}]$. This directly targets chunk-boundary inconsistency/jitter by stopping early when the model would otherwise continue with a materially different action trajectory than the one already being executed.

\paragraph{Decide-at-start rule: action-change magnitude (cheap prefix selection).}
This secondary ablation decides commit length at planning time from action-change magnitude using threshold $\delta$.

\paragraph{Decide-at-start rule: stochastic uncertainty (sample-based prefix selection).}
This secondary ablation decides commit length at planning time from sample dispersion using threshold $\epsilon$.

\paragraph{Learned commit head (auxiliary objective).}
To reduce manual threshold tuning, we optionally train a linear classifier on top of the observation encoder to predict a discrete commit class, supervised by the same longest-prefix rule applied to \emph{expert} actions in the dataset (expert action-change magnitude). At inference, the learned-commit mode queries this head once per replan to choose $\hat{m}\in[k_{\min},k_{\max}]$.

\paragraph{Comparator used in experiments: ACT-style temporal ensembling.}
In addition to queue-based execution, we evaluate an ACT-style temporal ensembling mode: at every timestep, the policy predicts a full chunk and deposits each predicted action into a per-timestep bin; the executed action is an exponentially weighted blend over all overlapping predictions for that timestep (older predictions weighted by $\exp(-m i)$, with $m$ configurable in evaluation). This serves as an ACT-aligned comparator to our adaptive queue/preemption method.

\paragraph{Train--deploy alignment for chunk CVAE.}
The chunk CVAE is trained with two reconstruction paths: (i) the standard posterior path using $q(z \mid o, a)$, and (ii) a deployment path using only prior-latent decoding (no future-action encoder input), weighted by a deployment-loss coefficient. We also align train and rollout action spaces by applying the same output squashing in reconstruction and inference. Offline evaluation reports both teacher-path loss and prior-$z{=}0$ action loss to better predict rollout behavior.

Thresholds $(\tau,\delta,\epsilon,w)$ and $k_{\max}$ are set in versioned experiment configurations for reproducibility.

\subsubsection{Overlap disagreement: score, preemption, and next commit length}
Let $c\in\mathbb{R}^{K\times A}$ be a candidate chunk and $q\in\mathbb{R}^{L\times A}$ the remaining FIFO queue (each row is one continuous action vector; $A$ is the action dimension). With overlap window $w$,
\[
m=\min\bigl(w,\,K,\,L\bigr),
\qquad
\text{score}=\frac{1}{m}\sum_{i=0}^{m-1}\bigl\|c_i-q_i\bigr\|_2 .
\]
If $\text{score}>\tau$, the executor \emph{preempts}: it replaces the queue with a \emph{prefix} of $c$ whose length $\hat{K}\in[k_{\min},k_{\max}]$ is computed from $\text{score}/\tau$ (ratio capped at $4$), linearly interpolating between $k_{\max}$ (small violation) and $k_{\min}$ (large violation), then clamping. If $\text{score}\le\tau$, no preempt occurs; the next environment step executes the first queued action from the FIFO queue. Initial planning under the overlap-disagreement policy loads $k_{\max}$ steps from the first chunk when the queue is empty.

\begin{algorithm}[H]
\caption{Overlap-disagreement execution (inference time)}
\begin{algorithmic}[1]
\State \textbf{Input:} policy $\pi$, thresholds $(\tau, w)$, bounds $(k_{\min}, k_{\max})$
\State Initialize empty queue $q$
\For{each timestep $t$}
  \If{$q$ is empty}
    \State $c \gets \pi(o_t)$; enqueue first $k_{\max}$ actions of $c$ into $q$
  \Else
    \State $c \gets \pi(o_t)$; $m \gets \min(w, |c|, |q|)$
    \State $\text{score} \gets \frac{1}{m}\sum_{i=0}^{m-1}\lVert c_i - q_i \rVert_2$
    \If{$\text{score} > \tau$}
      \State Compute $\hat{K}\in[k_{\min},k_{\max}]$ from capped ratio $\text{score}/\tau$
      \State Replace $q$ with first $\hat{K}$ actions of $c$
    \EndIf
  \EndIf
  \State Execute and pop the front action of $q$
\EndFor
\end{algorithmic}
\end{algorithm}

\subsubsection{Decide-at-start heuristics (ablations)}
\paragraph{Action-change magnitude.} From one deterministic chunk, we compute per-step $\ell_2$ action deltas and choose the longest prefix below $\delta$ (clamped to $[k_{\min},k_{\max}]$).
\paragraph{Stochastic uncertainty.} We draw $S$ chunk samples, compute per-step dispersion across samples, and choose the longest prefix below $\epsilon$.

\subsubsection{Dataset, training, and rollout evaluation}
\paragraph{Dataset.} We train from RoboMimic HDF5 demonstrations: each timestep stores low-dimensional observations and a continuous action vector. Lift, Can, and Square use \emph{separate} datasets and checkpoints (no multi-task mixing).
\paragraph{Training and offline validation.} Training is behavioral cloning: predict expert action chunks (or single actions for baselines) from normalized observations. Validation measures imitation loss on held-out demo timesteps; it does not use simulator rewards.
\paragraph{Rollouts.} Evaluation constructs a robosuite environment from metadata embedded in the dataset. For each execution mode, we run $20$ episodes by default; success rate is the fraction of episodes marked successful, and average return is the mean undiscounted sum of per-step rewards. Rewards are \emph{not} normalized by our code and are only comparable \emph{within} a task. Latent usage at rollout (sampled latent versus deterministic $z$) is selected per evaluation job, while execution mode is swept within that job.
\paragraph{Execution accounting.} Policy queries count forward passes through the policy (including batched stochastic forwards in uncertainty-based modes). Replans count executor plan refreshes (for example when preemption is triggered or the queue becomes empty), not every query. Commit statistics are computed from realized commit events during rollout; table entries such as \emph{Commit mean} report the mean realized commit length over those events.

\subsection{Baselines and comparators}

Table~\ref{tab:baselines} lists the imitation baselines and the ACT-style chunk policy we compare against (high-level descriptions; hyperparameters are fixed per task in our experiments).

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llp{6.2cm}@{}}
\toprule
Name & Role & Method summary \\
\midrule
BC-MLP & Single-step BC & Feedforward behavioral cloning baseline that predicts one action at a time from the current low-dimensional state. \\
ACT & Chunk transformer policy & Transformer chunking policy that predicts multi-step action sequences; execution mode (full chunk, receding horizon, dynamic, or temporal ensembling) is the main variable. \\
kNN-BC (VINN-style) & Nonparametric imitation & Observation-space nearest-neighbor action retrieval baseline (VINN-style lightweight variant \cite{vinn}). \\
BeT-style & Discrete behavior model & Discretized-action imitation baseline using behavior tokenization with a classifier over action bins \cite{bet}. \\
\bottomrule
\end{tabular}
\caption{Imitation baselines and ACT chunk-policy comparator evaluated in this work.}
\label{tab:baselines}
\end{table}

\noindent We evaluate BC-MLP, kNN-BC, BeT-style, and ACT variants under a shared simulator/metric protocol. BC-MLP is our in-project simple feedforward behavioral cloning baseline.

\subsection{Experiments}

We evaluate on RoboMimic Lift, Can, and Square with low-dimensional observations. We report episode success, return, horizon length, realized commit-length summaries ($\mathbb{E}[m]$, etc.), replans per episode, and policy forward-query counts (chunk forwards, stochastic multi-samples, and learned head forwards where applicable). Offline metrics include validation imitation loss (and commit CE when trained).

Execution modes for the chunk policy are full-chunk execution, receding-horizon execution, dynamic execution (with selectable rules), temporal ensembling, and learned-commit execution when the auxiliary head is trained. Single-step baselines (BC-MLP, kNN, BeT-style) are evaluated with receding-horizon execution so the policy class matches one-step control.

Dynamic rule in reported results: every \emph{dynamic} row and appendix entry uses the overlap-disagreement rule (mid-execution preemption). The decide-at-start heuristics (action-change magnitude and stochastic dispersion) are implemented but are not included in the reported rollout sweeps below.

For robustness, we optionally add Gaussian noise to the normalized observation vector at rollout time. We do not yet simulate object pose resets mid-episode.

For reproducibility, we fix Python/NumPy/Torch random seeds (default seed $42$) and retrain with multiple seeds (e.g.\ $0,1,2$) into separate run directories. Multi-seed summaries are reported for Lift, Can, and Square (Appendix Table~\ref{tab:seed_all_tasks}).

\section{Final Results}

Table~\ref{tab:act_main_cross_task} lists all four primary execution modes for each task under deterministic latent decoding. Our proposed execution rule is \textbf{Dynamic (overlap)} (mode label highlighted in bold in the table). Numbers are mean success rate and return over $20$ episodes per mode, with mean policy queries, replans, and realized commit statistics from the same logs. Returns are task-specific simulator rewards and are compared within each task (not averaged across tasks). Learned-commit rollouts use checkpoints trained with the auxiliary commit head; Table~\ref{tab:learned_commit} in the appendix reports those results. Lift learned offline metrics are: total loss $\approx 0.850$, action loss $\approx 0.222$, KL $\approx 6.16$, commit CE $\approx 1.232$, and prior-$z{=}0$ action loss $\approx 0.442$.

On Lift, our dynamic-overlap method reaches 1.00 success while reducing replans versus receding horizon (86.65 vs.\ 399.0). Can shows strong temporal ensembling and full-chunk performance under z0, while dynamic overlap underperforms in this snapshot (sampled-latent ablations partially recover dynamic; Appendix Table~\ref{tab:abl_sampled_all_tasks}). Square remains difficult across modes; receding horizon achieves slightly higher success than dynamic in this z0 table.
Learned-commit results (Appendix Table~\ref{tab:learned_commit}) show mixed transfer across tasks: Lift remains highly successful but with very short realized commits on average, Can attains moderate success with slightly longer commits, and Square remains difficult despite frequent replanning.

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llccccc@{}}
\toprule
Task & Mode & Success $\uparrow$ & Return $\uparrow$ & Queries $\downarrow$ & Repl./ep $\downarrow$ & Commit mean \\
\midrule
Lift & Temporal ens. & 1.00 & 76.10 & 400.0 & 0.00 & -- \\
Lift & Full chunk & 0.95 & 74.45 & 25.0 & 24.0 & 16.0 \\
Lift & Receding & 0.95 & 58.15 & 400.0 & 399.0 & 1.00 \\
Lift & \textbf{Dynamic (overlap)} & 1.00 & 60.00 & 400.0 & 86.65 & 4.56 \\
\midrule
Can & Temporal ens. & 1.00 & 289.15 & 400.0 & 0.00 & -- \\
Can & Full chunk & 1.00 & 268.15 & 25.0 & 24.0 & 16.0 \\
Can & Receding & 0.50 & 142.30 & 400.0 & 399.0 & 1.00 \\
Can & \textbf{Dynamic (overlap)} & 0.50 & 144.90 & 400.0 & 217.60 & 1.83 \\
\midrule
Square & Temporal ens. & 0.15 & 33.55 & 400.0 & 0.00 & -- \\
Square & Full chunk & 0.15 & 37.75 & 25.0 & 24.0 & 16.0 \\
Square & Receding & 0.20 & 41.70 & 400.0 & 399.0 & 1.00 \\
Square & \textbf{Dynamic (overlap)} & 0.15 & 42.05 & 400.0 & 184.20 & 2.16 \\
\bottomrule
\end{tabular}
\caption{ACT-Strong chunk policy across primary execution modes, with z0 rollout decoding and $20$ episodes per cell. Returns are task-specific simulator rewards and are not cross-task comparable.}
\label{tab:act_main_cross_task}
\end{table}

Table~\ref{tab:baseline_cross_task} reports one-step baseline results (receding-horizon eval for all baselines). Baselines are competitive on Lift, mixed on Can, and generally weak on Square, consistent with increased task difficulty in this low-dimensional setting. Our proposed method is \textbf{Dynamic (overlap)} in Table~\ref{tab:act_main_cross_task}; Table~\ref{tab:baseline_cross_task} contains only one-step baseline policies.

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}lccccc@{}}
\toprule
Task / baseline (receding) & Success $\uparrow$ & Return $\uparrow$ & Queries $\downarrow$ & Replans/ep $\downarrow$ & Commit mean \\
\midrule
Lift BC-MLP & 0.65 & 70.70 & 400.0 & 399.0 & 1.00 \\
Lift kNN-BC & 0.75 & 44.40 & 400.0 & 399.0 & 1.00 \\
Lift BeT-style & 0.75 & 127.30 & 400.0 & 399.0 & 1.00 \\
Can BC-MLP & 0.00 & 0.00 & 400.0 & 399.0 & 1.00 \\
Can kNN-BC & 0.30 & 58.65 & 400.0 & 399.0 & 1.00 \\
Can BeT-style & 0.30 & 33.75 & 400.0 & 399.0 & 1.00 \\
Square BC-MLP & 0.05 & 0.30 & 400.0 & 399.0 & 1.00 \\
Square kNN-BC & 0.35 & 43.00 & 400.0 & 399.0 & 1.00 \\
Square BeT-style & 0.05 & 5.60 & 400.0 & 399.0 & 1.00 \\
\bottomrule
\end{tabular}
\caption{Cross-task one-step imitation baselines under receding-horizon evaluation ($20$ episodes per cell).}
\label{tab:baseline_cross_task}
\end{table}

Across Lift multi-seed ACT-Strong runs (seeds 0/1/2), success remains $1.00 \pm 0.00$ (mean $\pm$ std). Can remains mixed and sensitive to execution/latent settings, while Square remains challenging for all tested policies in this configuration.
Additional all-task sampled-latent ablation snapshots and multi-seed aggregates for all modes are reported in Appendix Tables~\ref{tab:abl_sampled_all_tasks} and~\ref{tab:seed_all_tasks}.


\section{Discussion}
\paragraph{Hypothesis outcome.} Our central hypothesis is \emph{partially supported}. Adaptive execution is strongly supported on Lift, where dynamic commits match top success while reducing replans relative to receding horizon. On Can, behavior is mixed and sensitive to rollout setting (notably z0 versus sampled-latent ablations). On Square, all methods in our low-dimensional setup remain weak, indicating task difficulty dominates any single execution-rule choice under current training/data conditions.

\paragraph{Limitations.} Rollout evaluation requires the optional RoboMimic/MuJoCo stack (we use a separate rollout environment in practice, with Python version depending on MuJoCo compatibility). Lift success can be dominated by imitation quality; execution-mode differences can appear in commit statistics even when success is low. Large-scale vision-language imitation systems and full-scale VINN/BeT reproductions are outside this project scope.

\paragraph{Failure modes.} If disagreement/uncertainty thresholds are miscalibrated, the policy may replan too frequently (hurting stability) or too late (reducing reactivity). Threshold sweeps require editing experiment configuration and re-running evaluation.

\paragraph{Why Square remains difficult.} Square is contact-rich and tolerance-sensitive: small imitation or timing errors in peg insertion easily accumulate, so all policies we tested stay near chance-level success in the low-dimensional observation setting. When the underlying policy is weak, changing execution (receding vs.\ overlap dynamic vs.\ ensembling) mainly reshapes replanning cost and commit statistics rather than unlocking reliable success; stronger data, representations, or training signals likely matter more than a single heuristic execution rule.

\section{Conclusion}
We implemented inference-time adaptive commit-length execution for an ACT-style chunk policy, with heuristic rules, an optional learned commit head, imitation baselines (MLP BC, kNN, discrete-token BC), and shared rollout instrumentation. Final cross-task results support the proposal strongly on Lift, show mixed behavior on Can (including strong sampled-latent dynamic ablations), and reveal that Square is substantially harder under this low-dimensional setup.

Natural extensions include systematic threshold sweeps for overlap disagreement, reporting decide-at-start heuristics under the same evaluation matrix, adding seeds or confidence intervals beyond our three-seed summary, and moving from low-dimensional state to image observations where chunk-boundary effects may differ.

\section{Team Contributions}
In equal parts: Coding, Training, Inference. 
Charles: Ablations. 
Zuhayr: Analysis. 
Lawrence: Evaluation, benchmarking. 

\begin{itemize}
    \item Week 1: Architecture design, coding
    \item Week 2: Coding
    \item Week 3: Inference, ablations, evaluation
    \item Week 4: Analysis, writeup
    \item Week 5: Prepare Presentation
\end{itemize}


\appendix

\section{Ablation and seed summaries}

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llcc@{}}
\toprule
Task & Mode (sampled latent) & Success $\uparrow$ & Return $\uparrow$ \\
\midrule
Lift & Temporal ens. & 1.00 & 79.05 \\
Lift & Full chunk & 1.00 & 81.45 \\
Lift & Receding & 1.00 & 64.95 \\
Lift & Dynamic (overlap) & 1.00 & 63.80 \\
\midrule
Can & Temporal ens. & 1.00 & 289.75 \\
Can & Full chunk & 0.95 & 254.25 \\
Can & Receding & 0.80 & 230.60 \\
Can & Dynamic (overlap) & 0.95 & 274.90 \\
\midrule
Square & Temporal ens. & 0.45 & 118.00 \\
Square & Full chunk & 0.35 & 78.45 \\
Square & Receding & 0.25 & 50.90 \\
Square & Dynamic (overlap) & 0.30 & 70.95 \\
\bottomrule
\end{tabular}
\caption{All-task sampled-latent ablation snapshot for primary ACT modes ($20$ episodes per cell).}
\label{tab:abl_sampled_all_tasks}
\end{table}

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}llcc@{}}
\toprule
Task & Mode (3 seeds) & Success mean $\pm$ std & Return mean $\pm$ std \\
\midrule
Lift & Temporal ens. & 1.00 $\pm$ 0.00 & 225.35 $\pm$ 132.33 \\
Lift & Full chunk & 1.00 $\pm$ 0.00 & 208.83 $\pm$ 132.88 \\
Lift & Receding & 1.00 $\pm$ 0.00 & 177.73 $\pm$ 126.67 \\
Lift & Dynamic (overlap) & 1.00 $\pm$ 0.00 & 161.88 $\pm$ 130.14 \\
\midrule
Can & Temporal ens. & 0.97 $\pm$ 0.05 & 284.22 $\pm$ 16.50 \\
Can & Full chunk & 0.90 $\pm$ 0.04 & 255.43 $\pm$ 11.97 \\
Can & Receding & 0.50 $\pm$ 0.18 & 132.45 $\pm$ 45.63 \\
Can & Dynamic (overlap) & 0.65 $\pm$ 0.18 & 186.67 $\pm$ 51.66 \\
\midrule
Square & Temporal ens. & 0.25 $\pm$ 0.08 & 64.98 $\pm$ 19.78 \\
Square & Full chunk & 0.28 $\pm$ 0.13 & 73.80 $\pm$ 32.31 \\
Square & Receding & 0.12 $\pm$ 0.09 & 21.53 $\pm$ 23.58 \\
Square & Dynamic (overlap) & 0.15 $\pm$ 0.04 & 30.52 $\pm$ 15.33 \\
\bottomrule
\end{tabular}
\caption{All-task multi-seed summary ($3$ seeds: 0/1/2) for primary ACT modes.}
\label{tab:seed_all_tasks}
\end{table}

\section{Learned-commit appendix results}
This appendix reports learned-commit rollout metrics for each task under deterministic latent decoding.

\begin{table}[H]
\centering
\small
\begin{tabular}{@{}lccccc@{}}
\toprule
Task & Success $\uparrow$ & Return $\uparrow$ & Queries $\downarrow$ & Repl./ep $\downarrow$ & Commit mean \\
\midrule
Lift & 1.00 & 44.65 & 755.0 & 376.50 & 1.06 \\
Can & 0.75 & 224.40 & 630.4 & 314.20 & 1.27 \\
Square & 0.15 & 40.60 & 746.3 & 372.15 & 1.07 \\
\bottomrule
\end{tabular}
\caption{Learned-commit results under z0 decoding, with $20$ episodes per task.}
\label{tab:learned_commit}
\end{table}

\section{AI Usage Declaration}

We used AI assistants (Cursor/ChatGPT) as coding and writing support tools: to refine prose, suggest command-line workflows, and help diagnose implementation issues. All experiment design decisions, code integration, run execution, result validation, and final report content were reviewed and finalized by the authors.
\noindent\textbf{Code repository:} \url{https://github.com/Jojoz23/actdyn}



\bibliographystyle{iclr2025_conference}
\bibliography{refs}





\end{document}
